---
title: 'Assignment 5'
author: "Alia Ajina and Sophia Leiker"
date: "March 4, 2023"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Loading in required packages
library(tidyverse)
library(here)
library(testthat)
library(purrr)
library(devtools)
library(ggplot2)
library(tidyr)
library(cowplot)
library(kableExtra)
```

# Assignment 5
In this assignment, we created a function to calculate lake level based on four initial lake levels at Huntington Lake in the Sierra Nevadas. These lake levels are: 100,000 gallons, 120,000 gallons, 140,000 gallons, and 160,000 gallons. Additionally we assumed that outflow from the lake stays even at 10 gallons/day and evaporation stays equal at 1 gallon/day. This equation is function for one single day. Precipitation data (4 values) was created using a random sample generator with a mean of 10 and a standard deviation of 2, these values are in gallons.

## Sourcing Our Function
```{r}
# Sourcing our equation
source(here("R", "lake_equation.R"))
```

## Creating Dataset to run our function on
```{r }
# Creating our initial lake level dataset
initial_level <- c(100000,120000,140000,160000)

# Creating our random sample of daily precipitation inputs
nsample = 4
precip_data = rnorm(mean = 10, sd = 2, n=nsample)
precip_results = data.frame(precip_data=precip_data)

# Creating inflow and outflow 
outflow = c(10,10,10,10)
evap = c(1,1,1,1)

# Combining into 1 dataframe
possible_lake <- data.frame(initial_level, outflow, evap)

# Checking the outputs
# head(possible_lake)
```

## Checking that our function works
```{r, by hand checking that our function works}

# Checking that our function works
# lake_func(100,10)

```


## Using Pmap
```{r}

# Using pmap to run our function through our dataset
lake_output = pmap(list(initial_level = initial_level, outflow = outflow, evap = evap), lake_func, precip_data = precip_data)

# Checking our output from pmap
# print(lake_output)

# Putting our output into a datafram
lake_data = as.data.frame(lake_output)
# print(lake_data)

# Renaming our columns
col_names = c("Lake 1 Precip 1: Lake Volume (gal)", "Lake 1 Precip 1: Total Change (gal)", "Lake 2 Precip 2: Lake Volume (gal)", "Lake 2 Precip 2: Total Change (gal)", "Lake 3 Precip 3: Lake Volume (gal)", "Lake 3 Precip 3: Total Change (gal)", "Lake 4 Precip 4: Lake Volume (gal)", "Lake 4 Precip 4: Total Change (gal)")

colnames(lake_data) = col_names

# Making Table Output
kable_table <- kable(lake_data, caption = "Results using Pmap")
kable_table %>% 
  kable_styling(full_width = FALSE, 
                bootstrap_options = "bordered",
                position = "center")

```

## Using for loop
```{r}

# Using a for loop to run our function through our dataset and creating a dataframe
for_loop_out = as.data.frame(matrix(nrow=length(precip_data), ncol=nrow(possible_lake)))

for (i in 1:nrow(possible_lake)) {
  result = lake_func(initial_level=possible_lake$initial_level[i], outflow = possible_lake$outflow[i],evap = possible_lake$evap[i], precip_data = precip_data)
  for_loop_out[,2*i-1] = result[[1]]   # "lake_vol" to odd-numbered columns
  for_loop_out[,2*i] = result[[2]]     # "total_change" to even-numbered columns
}

# Checking our output
# print(for_loop_out)

# Updating Column Names
colnames(for_loop_out) = col_names

# Checking our output
# print(for_loop_out)

# Making Table Output
kable_table <- kable(for_loop_out, caption = "Results using For Loop")
kable_table %>% 
  kable_styling(full_width = FALSE, 
                bootstrap_options = "bordered",
                position = "center")
```

## Making a Graph
```{r}
# Pulling apart dataframes
vol_graphs <- for_loop_out[, c(1, 3, 5, 7)]
change_graphs <- for_loop_out[, c(2, 4, 6, 8)]

rownames(change_graphs) <- c("Precip 1", "Precip 2", "Precip 3", "Precip 4")

# Creating 4 plots for each precipitation
plot1 <- ggplot(data = change_graphs, aes(x = rownames(change_graphs), y = change_graphs[, 1], fill = rownames(change_graphs))) +
  geom_col() +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  ggtitle("Lake 1: Total Change for Each Precipitation Input") + 
  theme(title = element_text(size=8)) +
  ylab("Changed Volume (gal)") +
  xlab("Input Precipitation Levels") +
  guides(fill = F)
  
plot2 <- ggplot(data = change_graphs, aes(x = rownames(change_graphs), y = change_graphs[, 2], fill = rownames(change_graphs))) +
  geom_col() +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  ggtitle("Lake 2: Total Change for Each Precipitation Input") + 
  theme(title = element_text(size=8)) +
  ylab("Changed Volume (gal)") +
  xlab("Input Precipitation Levels") +
  guides(fill = F)
  
plot3 <- ggplot(data = change_graphs, aes(x = rownames(change_graphs), y = change_graphs[, 3], fill = rownames(change_graphs))) +
  geom_col() +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  ggtitle("Lake 3: Total Change for Each Precipitation Input") + 
  theme(title = element_text(size=8)) +
  ylab("Changed Volume (gal)") +
  xlab("Input Precipitation Levels") +
  guides(fill = F)
  
plot4 <- ggplot(data = change_graphs, aes(x = rownames(change_graphs), y = change_graphs[, 4], fill = rownames(change_graphs))) +
  geom_col() +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  ggtitle("Lake 4: Total Change for Each Precipitation Input") + 
  theme(title = element_text(size=8)) +
  ylab("Changed Volume (gal)") +
  xlab("Input Precipitation Levels") +
  guides(fill = F)

# Arrange plots in a grid
plot_grid(plot1, plot2, plot3, plot4, ncol = 2)

```

